---
title: "IMDB Ratings"
author: "Kaiyuan Li"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

knitr::opts_chunk$set(
  fig.width=8, 
  fig.height=6,
  fig.align = "center"
)
```

```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(dplyr)
library(ggplot2)
```

# Analysis of movies- IMDB dataset

```{r,load_movies, warning=FALSE, message=FALSE}
# Assign IMDB data to variable
movies <- read_csv(here::here("data", "movies.csv"))

```

## Use your data import, inspection, and cleaning skills to answer the following:

-   Are there any missing values (NAs)? Are all entries distinct or are there duplicate entries?

There are no missing values according to the skim function. Nevertheless, we notice some duplicate values, for example "Nightmare on Elm Street" or "Alice in Wonderland". We have to be careful to look at more than just the title in identifying duplicates, as some movies could have the same name, but be released twice on different dates or have a different director.

```{r,Skim_data}
# Skim IMDB data
skim(movies)

```

-   Produce a table with the count of movies by genre, ranked in descending order

    ```{r,count_of_movies}
    movies%>%
      count(sort = TRUE,genre)
    ```

-   Produce a table with the average gross earning and budget (`gross` and `budget`) by genre. Calculate a variable `return_on_budget` which shows how many \$ did a movie make at the box office for each \$ of its budget. Ranked genres by this `return_on_budget` in descending order

    ```{r,avg_gross_budget}
    # Create table with the average gross earning and budget
    avg_genre = movies%>%
      group_by(genre) %>% 
      summarize(avg_earning = sum(gross)/count(genre),
                avg_budget = sum(budget)/count(genre)) # Create 2 columns to store the average earning and budget
    avg_genre%>% 
      mutate(return_on_budget = avg_earning/avg_budget) %>% # The return is just the earning/budget
      arrange(desc(return_on_budget))
    ```

-   Produce a table that shows the top 15 directors who have created the highest gross revenue in the box office. Don't just show the total gross amount, but also the mean, median, and standard deviation per director.

    ```{r,top_15}
    # Calculate summary statistics for top 15 directors 
    Top_directors = movies%>%
      group_by(director) %>% 
      summarise(sum_gross = sum(gross),
                mean_gross = mean(gross),
                median_gross = median(gross),
                SD_gross = sd(gross))
    # Choose the top 15 with highest gross earnings
    Top_directors%>%
      slice_max(sum_gross,n = 15)
    ```

-   Finally, ratings. Produce a table that describes how ratings are distributed by genre. We don't want just the mean, but also, min, max, median, SD and some kind of a histogram or density graph that visually shows how ratings are distributed.

    ```{r,rating}
    # Calculate summary statistics for ratings by genre
    Ratings_genre = movies%>%
      group_by(genre)%>%
      summarise(mean_ratings = mean(rating),
                min_rating = min(rating),
                max_rating = max(rating),
                median_rating = median(rating),
                SD_rating = sd(rating))
    Ratings_genre

    # Plot ratings by genre
    ggplot(movies,aes(x=rating)) +
      geom_density() +
      labs(title="There are few completely unpopular movies with a rating of less than 5",subtitle = "Density plot of movie ratings on IMDB",x = "Rating", y = "Density")+
      theme_bw()
    ```

## Use `ggplot` to answer the following

-   Examine the relationship between `gross` and `cast_facebook_likes`. Produce a scatterplot and write one sentence discussing whether the number of facebook likes that the cast has received is likely to be a good predictor of how much money a movie will make at the box office. What variable are you going to map to the Y- and X- axes?

While there seems to be a minor correlation between the amount of cast facebook likes and the money a movie makes, the relationship is not strong enough to make it a good predictor of a movie's success.

```{r, gross_on_fblikes}
# Map revenue vs Facebook likes
movies%>%
ggplot(aes(x=cast_facebook_likes, y = gross)) +
  geom_point()+scale_x_log10()+geom_smooth(method = "lm", se = FALSE) +
  labs(title="Cast Facebook likes do not seem to be a reliable predictor of movie success",subtitle = "Scatterplot of number of cast facebook likes and movie gross revenue",x = "Number of Cast Facebook Likes", y = "Gross Revenue")+
  theme_bw()
```

-   Examine the relationship between `gross` and `budget`. Produce a scatterplot and write one sentence discussing whether budget is likely to be a good predictor of how much money a movie will make at the box office.

While a budget is surely not a guarantor of movie success , the fitted line has a positive slope and therefore implies that with rising budget comes rising gross revenue.

```{r, gross_on_budget}
 #Map revenue vs budget
movies%>%
  ggplot(aes(x=budget,y=gross))+
  geom_point()+geom_smooth(method = "lm", se = FALSE) +
  labs(title="A higher budget seems to positively affect movie gross revenue",subtitle = "Scatterplot of movie budget and movie gross revenue",x = "Budget", y = "Gross Revenue")+
  theme_bw()

```

-   Examine the relationship between `gross` and `rating`. Produce a scatterplot, faceted by `genre` and discuss whether IMDB ratings are likely to be a good predictor of how much money a movie will make at the box office. Is there anything strange in this dataset?

Generally, higher ratings indicates higher gross earnings for all genres for which we have a significant amount of data. However, we can also see that it is possible for movies to have a good rating while not making a lot of money. This most likely concern the likes of indie movies, that receive strong support but never make it into the pop culture.

There are some interesting anomalies in the data in the form of extreme values. Drama movies usually fall within the same range of gross revenues, however "Titanic" by James Cameron reports much higher values than the rest. Coincidentally, James Cameron is also the director for highest grossing movie in the data "Avatar". It is also noticeable that there are no observable outliers for genres like comedy or adventure, even though there are several movies of the genre in the data. Apparently these categories do not display the required parameters to polarize the nation.

```{r, gross_on_rating}
# Map revenue vs rating, faceted by genre
movies%>%
  ggplot(aes(x=rating, y = gross,color=genre))+
  geom_point()+facet_wrap(~genre)+
  labs(title="A higher rating seems to be correlated to higher gross revenues",subtitle = "Faceted scatterplot of IMDB rating and movie gross revenue",x = "IMDB Rating", y = "Gross Revenue")
```

# IMDB ratings: Differences between directors

Recall the IMBD ratings data. I would like you to explore whether the mean IMDB rating for Steven Spielberg and Tim Burton are the same or not. I have already calculated the confidence intervals for the mean ratings of these two directors and as you can see they overlap.

First, I would like you to reproduce this graph. You may find `geom_errorbar()` and `geom_rect()` useful.

In addition, you will run a hypothesis test. You should use both the `t.test` command and the `infer` package to simulate from a null distribution, where you assume zero difference between the two.

> Before anything, write down the null and alternative hypotheses, as well as the resulting test statistic and the associated t-stat or p-value. At the end of the day, what do you conclude?

**H0** = There is no difference in the average IMDB rating of Tim Burton and Steven Spielberg movies **H1** = There is a difference in the average IMDB rating of Tim Burton and Steven Spielberg movies

You can load the data and examine its structure

```{r load-movies-data}

# Load and glimpse movie data
movies <- read_csv(here::here("data", "movies.csv"))
glimpse(movies)

```

```{r director_plot}

# Create list to filter for directors of interest
direc <- c("Tim Burton","Steven Spielberg")
library(ggrepel)

movies %>% 
  # filter and group for directors
  filter(director %in% direc) %>% 
  group_by(director) %>% 
  # Calculate summary statistics
  summarize(mean_rating = mean(rating),
            sd_rating = sd(rating),
            count = n(),
            t_critical = qt(0.975, count-1)) %>%
  mutate(x_min = mean_rating - t_critical*sd_rating/sqrt(count), 
         x_max = mean_rating + t_critical*sd_rating/sqrt(count) ) %>% 
  # Create confidence interval plot for directors
  ggplot(aes(x = mean_rating,y=director, colour = director,
         label = round(mean_rating,2), size = 10))+
  geom_pointrange(aes(xmin=x_min,xmax=x_max),size=1)+
  geom_errorbar(aes(xmin = x_min, xmax = x_max),
                width = 0.1, size = 1.2) +
  # Creating grey angle to mark overlap between intervals
  geom_rect(aes(xmin = x_min[1], xmax = x_max[2], 
            ymin = 0, ymax = 3),
            alpha = 1/8,color = NA) +
  # Add labels of margins of the confidence intervals
  geom_text(aes(label = round(x_min,2), x=x_min),
            vjust = -1, size = 3.5,
            color = "black") +
  geom_text(aes(label = round(x_max,2), x=x_max), 
            vjust = -1, size = 3.5,
            color = "black") +
  geom_text(aes(label = round(mean_rating,2), x=mean_rating),
            vjust = -1, size = 5,
            color = "black") +
  # Format plot
  scale_y_discrete(limits = c("Tim Burton","Steven Spielberg"))+
  theme_bw()+
  theme(legend.position = "none")+
  labs(title = "Do Spielberg and Burton have the same mean IMDB ratings?",
       subtitle = "95% confidence intervals overlap",
       x = "Mean IMDB Rating", 
       y = NULL) +
  theme(plot.title = element_text(face="bold"))
  
direc_data <- movies %>% 
  filter(director %in% direc) %>% 
  group_by(director) %>% 
  summarize(rating, n = n())

# Run t-test   
t.test(rating ~ director, data = direc_data )
```

As we can see from the two sample t-test, we reject the null hypothesis as 0 does not lie within the confidence interval. We can therefore say with 95% confidence that there is a signifcant difference in the average rating of Steven Spielberg and Tim Burton films. The graph confirms this, as we see only a very small and seemingly insignificant overlap between the confidence intervals. Specifically, we also see that the difference stems from Steven Spielberg having a **higher** average rating. Nevertheless, it should be mentioned that the sample size for both directors is rather low, with a correspondingly high t_critical. As more movies come out and the ratings do not drastically deviate from the current average, we might expect to not see any more overlap of the confidence intervals in the future.

Personally, we enjoy Beetlejuice just as much as Jurassic Park.
